{\steve}: A hierarchical Bayesian model for Supernova Cosmology

### 2018-12-19 papers/methods/figures/tikz.tex
Updating paper for referees comments

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.8cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.8cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
	\node[det, above=of zobs] (z) {$z$};
	\node[det, above=of z] (mu) {$\mu$};
	\node[latent, right=4.33cm of mu] (MBobs) {$M_B$};
	% Survey nodes
	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
	% Global nodes
	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
	% Connect the nodes
	\edge {sum} {obssum};
	\edge {z} {zobs};
	\edge {z,cosmo} {mu};
	\edge {mu, sum, mobs, pop1, ab, pobs, sf} {MBobs};
	% Plates
	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

>>>>

% Use vector fonts, so it zooms properly in on-screen viewing software
% Don't change these lines unless you know what you are doing
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	% Including figure files
\usepackage{amsmath}	% Advanced maths commands
\usepackage{amssymb}	% Extra maths symbols
\usepackage{tikz}
\DeclareMathAlphabet{\mathcal}{OMS}{cmsy}{m}{n}
\usetikzlibrary{bayesnet}
\usetikzlibrary{external}
\tikzexternalize % activate!
\tikzset{external/force remake}
\begin{document}
	\begin{tikzpicture}
	% Obs nodes
	\node[obs]  (obssum) {$\hat{m}_B, \hat{x}_1, \hat{c}, \mathcal{C_\eta}^{-1}$};
	\node[obs, left=0.8cm of obssum]  (zobs) {$\hat{z}$};
	\node[obs, right=0.6cm of obssum]  (mobs) {$\hat{m}$};
	\node[obs, right=0.6cm of mobs]  (pobs) {$\hat{p}$};
	% Latent nodes
	\node[latent, above=of obssum] (sum) {$m_B, x_1, c$};
	\node[det, above=of zobs] (z) {$z$};	
	\node[det, above=of z] (mu) {$\mu$};
	\node[det, right=4.33cm of mu] (MBobs) {$M_B$};
	% Survey nodes
	\node[latent, above=of sum, yshift=1cm,text width=2.7cm, align=center] (pop1) {$\langle x_1^i \rangle, \langle c^i \rangle,\sigma_{M_B}, $\\$ \sigma_{x_1}, \sigma_c, \alpha_c, \kappa_0, \kappa_1$ };
	\node[latent, right=1.5cm of pop1, align=center] (sf) {$\delta S$};
	% Global nodes
	\node[latent, above=2cm of mu]  (cosmo) {$\Omega_m, w$};
	\node[latent, above=1.7cm of MBobs, text width=3cm, align=center]  (ab) {$\langle M_B \rangle, \alpha, \beta$, $\delta(0)$, \\$\delta(\infty)$, $\delta \mathcal{Z}_i$};
	% Connect the nodes
	\edge {sum} {obssum};
	\edge {z} {zobs};
	\edge {z,cosmo} {mu};
	\edge {mu, sum, pop1, ab, sf, pobs, mobs} {MBobs};	
	% Plates
	\plated[thick] {sn} {(obssum)(zobs)(mobs)(pobs)(sum)(z)(mu)} {$\forall\ \rm{SN}$} ;
	\plated[thick] {survey} {(sn)(pop1)} {\\$\forall\ \rm{Survey}$} ;
	\end{tikzpicture}
\end{document}

### 2018-12-19 papers/methods/paper.tex
Updating for arxiv upload
Updating paper for referees comments

%\usepackage{lineno}
%\linenumbers
\renewcommand{\arraystretch}{1.2}
\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement) but instead uncertainty on distance modulus. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.

@@

Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space. To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.

In addition to absolute magnitude, the underlying supernova population is also characterised by distributions in color and stretch. We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey. For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution. We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude. Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the color and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment. We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ that allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. We sample $\delta_c$ in log space for efficiency in sampling close to zero. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors.

@@

We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM. $\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$. For calculating the distance modulus, we fix $H_0 = 70 \kmsmpc $. If the Hubble constant has a different value, the absolute magnitude is $M_B + 5\log(H_0/70 \kmsmpc )$ with the other cosmological parameters unaffected.

>>>>

%\usepackage{lineno}
%\linenumbers
\renewcommand{\arraystretch}{1.2}
%\graphicspath{ {./figures/} }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

As we are focused on the spectroscopically confirmed supernovae for this iteration of the method, we assume the observed redshift $\hat{z}$ is the true redshift $z$ such that $P(\hat{z}|z) = \delta(\hat{z} - z)$. Potential sources of redshift error (such as peculiar velocities) are taken into account not via uncertainty on redshift (which is technically challenging to implement {\blue as varying redshifts introduce computational complexity in computing the distance modulus integral}) but instead uncertainty on distance modulus. Similarly, we take the mass probability estimate $\hat{m}$ as correct, and do not model a latent variable to represent uncertainty on the probability estimate. One of the strengths of {\steve} (and the {\rubin} analysis) is that for future data sets where supernovae have been classified photometrically, and we expect some misclassification and misidentification of the host galaxies, these misclassifications can naturally be modelled and taken into account by introducing additional populations that supernovae have a non-zero probability of belonging to.

@@

Unlike many previous formalisms which utilise $M_B$ as a singular number and model magnitude scatter on the apparent magnitude $m_B$, we incorporate this scatter into the underlying rest-frame population by having a population in absolute magnitude space. {\blue This is mathematically equivalent, however allows us to model the underlying population and intrinsic scatter distinctly.} To denote this difference, we refer to the mean of our absolute magnitude population with $\langle M_B \rangle$.

In addition to absolute magnitude, the underlying supernova population is also characterised by distributions in color and stretch. We follow the prior work of {\rubin} and model the color population as an independent redshift-dependent skew normal distribution for each survey. For the stretch population, we adopt a redshift-dependent normal distribution, and magnitude dispersion is modelled as a normal distribution. We also tested a skew-normal approach for these parameters, reverting to the normal distributions as they are computationally easier to evaluate and we found no reduction in cosmological bias with the skew-normal distributions for stretch and magnitude. Following {\rubin} we allow the mean color and stretch to vary over redshift, anchoring four equally spaced redshift nodes spanning the redshift range of each survey, linearly interpolating between the nodes to determine the mean stretch and color for a given redshift. These nodes are represented as $\langle x_1^i \rangle$ and $\langle c^i \rangle$. Both the color and stretch means are modelled with normal priors. Initial versions of our model adopted a fully covariant multivariate skew normal (with skewness set to zero only for the magnitude component) to capture correlations between $m_B$ and $c$, however pathological fitting complications required us to simplify our treatment. We parametrise the color skewness $\alpha_c$ by sampling $\delta_c = \alpha_c / \sqrt{1 + \alpha_c^2}$ which itself is given a uniform prior $\mathcal{U}(0,0.98)$ that allows $\alpha_c$ to span positive values in the realm of physical plausibility as determined from constraints in \citet{Scolnic2016}. We sample $\delta_c$ in log space for efficiency in sampling close to zero. The width of the population, represented by the vector $\lbrace \sigma_{M_B}, \sigma_{x_1}, \sigma_c \rbrace$ is subject to Cauchy priors with {\blue mean 0 and width 1}.

@@

We formulate our model with three different cosmological parameterisations; Flat $\Lambda$CDM, Flat $w$CDM, and standard $\Lambda$CDM {\blue (without a flatness prior)}. $\Omega_m$ is given the prior $\mathcal{U}(0.05, 0.99)$, $\Omega_\Lambda$ was treated with $\mathcal{U}(0, 1.5)$ and the equation of state $w$ was similarly set to a flat prior $\mathcal{U}(-0.4, -2.0)$. For calculating the distance modulus, we fix $H_0 = 70 \kmsmpc $. If the Hubble constant has a different value, the absolute magnitude is $M_B + 5\log(H_0/70 \kmsmpc )$ with the other cosmological parameters unaffected.
